Guided Proofreading of Automatic Segmentations for Connectomics

### 2016-11-13 paper/guidedproofreading_review.tex
Changed Paper ID
Paper rewrite.

\def\cvprPaperID{****} % *** Enter the CVPR Paper ID here
\def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}

@@

%%%%%%%%% TITLE
\title{Guided Proofreading of Automatic Segmentations in Connectomics}

@@

%%%%%%%%% ABSTRACT
\begin{abstract}
Automatic cell image segmentation methods in connectomics produce merge and split errors, which require correction through proofreading. Previous research has identified the visual search for these errors as the bottleneck in interactive proofreading. To aid error correction, we develop two classifiers to recommend candidate merge and split errors and their corrections to the user. These classifiers are informed by training a convolutional neural network with known errors in automatic segmentations against expert-labeled ground truth. Our classifiers detect potentially-erroneous regions by considering a large context region around a segmentation boundary. Corrections can then be performed as yes/no decisions 
resulting in faster correction times than previous methods. We evaluate our approach on connectomics datasets of different species and compare correction performance of novice and expert users against different existing systems. We report significant improvements on pure automatic and pure manual proofreading.
\end{abstract}

@@

>>>>

\def\cvprPaperID{0947} % *** Enter the CVPR Paper ID here
\def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}

@@

%%%%%%%%% TITLE
\title{Guided Proofreading of Automatic Segmentations for Connectomics}

@@

%%%%%%%%% ABSTRACT
\begin{abstract}
Automatic cell image segmentation methods in connectomics produce merge and split errors, which require correction through proofreading. Previous research has identified the visual search for these errors as the bottleneck in interactive proofreading. To aid error correction, we develop two classifiers to recommend candidate merge and split errors and their corrections to the user. These classifiers are informed by training a convolutional neural network with known errors in automatic segmentations against expert-labeled ground truth. Our classifiers detect potentially-erroneous regions by considering a large context region around a segmentation boundary. Corrections can then be performed as yes/no decisions 
resulting in faster correction times than previous methods. We evaluate our approach on connectomics datasets of different species and compare correction performance of novice and expert users against different existing systems. We report significant improvements compared to pure automatic and pure manual proofreading.
\end{abstract}

@@

\input{results}

### 2016-11-13 paper/introduction.tex
Paper rewrite.

All automatic methods make errors, and we are left with large data which needs \emph{proofreading} by humans. This crucial task serves two purposes: 1) to correct errors in the segmentation, and 2) to provide a large body of labeled data to train better automatic segmentation methods. Recent proofreading tools provide intuitive user interfaces to browse segmentation data in 2D and 3D and to identify and manually correct errors ~\cite{markus_proofreading,raveler,mojo2,haehn_dojo_2014}. Many kinds of errors exist, such as inaccurate boundaries, but the most common are \emph{split errors}, where a single segment is labeled as two, and \emph{merge errors}, where two segments are labeled as one (Fig.~\ref{fig:merge_and_slit_errors}). With user interaction, split errors can be joined, and the missing boundary in a merge error can be defined with manually-seeded watersheds~\cite{haehn_dojo_2014}. However, even with semi-automatic correction tools, the visual inspection to find errors in the first place takes the majority of the time.

@@

Our goal is to add automatic detection of split and merge errors to proofreading tools. Instead of the user visually inspecting the whole data volume carefully to spot errors, we design automatic classifiers that detect split and merge errors in 2D segmentations. Then, a proofreading tool can recommend regions with a high probability of an error to the user, and suggest corrections to accept or reject. We call this process \textit{guided proofreading} (GP).

As our main contribution, we introduce classifiers to detect merge- and split errors based on a convolutional neural network (CNN). We believe that this is the first time that deep learning is applied to the task of proofreading --- especially in the field of connectomics. Our classifiers work on top of any existing automatic segmentation method to find potential errors and suggest corrections. Given a membrane segmentation from a fast automatic method, our classifiers operate on the boundaries of whole cell regions. Compared to techniques that must analyze every input pixel, we reduce the data analysis to the boundaries only. Explicitly, we train a CNN to detect only split errors. We then reuse the same network to also detect merge errors by generating possible boundaries within a cell and inverting the split error score. Correction suggestion for both types of errors is then trivial and our technique therefor reduces the proofreading operation to simple yes/no decisions.

We further propose a greedy algorithm to perform proofreading automatically and measure performance as the adapted Rand error which is a common metric for segmentation comparison~\cite{RAND}. 
Our system is integrated into an existing proofreading workflow for large connectomics data. For this, we also explore an active label suggestion approach in addition to the ranking obtained by guided proofreading. We quantitatively validate automatic and human-driven variations of guided proofreading on five different real-world connectomics datasets of mouse as well as fruitfly (drosophila) brain. To study the performance of novice and expert proofreaders, we perform a between-subjects experiment and ask participants to proofread a publicly available dataset. For comparison, we establish two baselines: a recently published fully interactive proofreading tool named \textit{Dojo} by Haehn~\etal~\cite{haehn_dojo_2014} and semi-automatic \textit{focused proofreading} (FP) approach by Plaza~\cite{focused_proofreading}. In all experiments, we significantly outperform both interactive proofreading as well as Plaza's method. As a consequence, we are able to provide tools to proofread segmentations more efficiently, and so better tackle large volumes of connectomics imagery.

>>>>

All automatic methods make errors, and we are left with large data which needs \emph{proofreading} by humans. This crucial task serves two purposes: 1) to correct errors in the segmentation, and 2) to provide a large body of labeled data to train better automatic segmentation methods. Recent proofreading tools provide intuitive user interfaces to browse segmentation data in 2D and 3D and to identify and manually correct errors ~\cite{markus_proofreading,raveler,mojo2,haehn_dojo_2014}. Many kinds of errors exist, such as inaccurate boundaries, but the most common are \emph{split errors}, where a single segment is labeled as two, and \emph{merge errors}, where two segments are labeled as one (Fig.~\ref{fig:merge_and_slit_errors}). With user interaction, split errors can be joined, and the missing boundary in a merge error can be defined with manually-seeded watersheds~\cite{haehn_dojo_2014}. However, even with semi-automatic correction tools, the visual inspection to find errors takes the majority of the time~\cite{proofreading_bottleneck}.

@@

Our goal is to add automatic detection of split and merge errors to proofreading tools. We design automatic classifiers that detect split and merge errors in 2D segmentations so the user does not need to visually inspect the whole data volume to spot errors. A proofreading tool then recommends regions with a high probability of an error to the user, and suggest corrections to accept or reject. We call this process \textit{guided proofreading} (GP).

As our main contribution, we introduce classifiers to detect merge- and split errors based on a convolutional neural network (CNN). We believe that this is the first time that deep learning is applied to the task of proofreading. Our classifiers work on top of any existing automatic segmentation method to find potential errors and suggest corrections. Given a membrane segmentation from a fast automatic method, our classifiers operate on the boundaries of whole cell regions. Compared to techniques that must analyze every input pixel, we reduce the data analysis to the boundaries only. First, we train a CNN to detect only split errors. The output of this network is a probability whether a boundary between two segments is valid or not. We then reuse the same network to also detect merge errors by generating possible boundaries within a cell and inverting the split error score. We create corrections for both types of errors which can be accepted or rejected. This reduces the proofreading operation to simple yes/no decisions.

We further propose a greedy algorithm to perform proofreading automatically. Possible erroneous regions are sorted by their score and the algorithm sequentially corrects merge errors and split errors until a configurable threshold is reached. If ground truth data is available, we can use a selection oracle to drive this algorithm. The oracle only accepts corrections which improve the automatic segmentation.

This way,

 and measure performance as the adapted Rand error which is a common metric for segmentation comparison~\cite{RAND}. 
Our system is integrated into an existing proofreading workflow for large connectomics data. For this, we also explore an active label suggestion approach in addition to the ranking obtained by guided proofreading. We quantitatively validate automatic and human-driven variations of guided proofreading on five different real-world connectomics datasets of mouse as well as fruitfly (drosophila) brain. To study the performance of novice and expert proofreaders, we perform a between-subjects experiment and ask participants to proofread a publicly available dataset. For comparison, we establish two baselines: a recently published fully interactive proofreading tool named \textit{Dojo} by Haehn~\etal~\cite{haehn_dojo_2014} and semi-automatic \textit{focused proofreading} (FP) approach by Plaza~\cite{focused_proofreading}. In all experiments, we significantly outperform both interactive proofreading as well as Plaza's method.

As a consequence, we are able to provide tools to proofread segmentations more efficiently, and so better tackle large volumes of connectomics imagery.

### 2016-11-13 paper/related_work.tex
Paper rewrite.

Proofreading is performed on top of an existing automatic segmentation. First, this section gives an overview of state-of-the-art automatic segmentation methods. Then, we discuss interactive proofreading tools and recent developments in computer-aided proofreading systems.

@@

NeuroProof \cite{neuroproof2013} tries to decrease error rates with interactive learning of agglomeration of over-segmentations of images, based on a random forest classifier. Vazquez-Reina \etal~\cite{amelio_segmentation} propose automatic 3D segmentation by taking whole EM volumes into account rather than a per section approach, then solving a fusion problem with a global context. Kaynig \etal~\cite{kaynig10} propose a random forest classifier coupled with an anisotropic smoothing prior in a conditional random field framework with 3D segment fusion. It is also possible to learn segmentation classification features directly from images with CNNs. Ronneberger \etal~\cite{RonnebergerFB15} use a contracting/expanding CNN path architecture to enable precise boundary localization with small amounts of training data. Lee \etal~\cite{lee2015recursive} recursively train very deep networks with 2D and 3D filters to detect boundaries. Bogovic \etal~\cite{BogovicHJ13} learn 3D features, and show even that unsupervised learning can produce better features than hand-designs. Our work was inspired by this paper and we extend the features reported by Bogovic \etal for our guided proofreading classifiers as described in section \ref{sec:methods}.
These approaches make good progress; however, in general, proofreading is required to improve them through generating more ground-truth segmentations.

\textbf{Interactive Proofreading.} While proofreading is very time consuming, it is fairly easy for humans to perform corrections through splitting and merging segments. One way to perform such corrections is by using expert tools such as Raveler introduced by Chklovskii~\etal~\cite{chklovskii2010, raveler}. This software offers many parameters for tweaking the proofreading process. Created in 2010, Raveler is still used today by professional full-time proofreaders and many similar systems exist as stand-alone products or plugins to existing visualization system,~\eg V3D~\cite{proofreading_bottleneck} or AVIZO~\cite{markus_proofreading}. In contrast to these expert tools, recent works attack the problem of proofreading massive datasets by novices through crowd-sourcing~\cite{saalfeld09,anderson2011,Giuly2013DP2}. A very popular platform is EyeWire presented by Kim \etal~\cite{eyewire_nature}. EyeWire is set up as an online game and participants earn virtual rewards for merging oversegmented labeling to reconstruct the retina cells.
A range of proofreading tools exist in-between expert systems and online games such as Mojo and \textit{Dojo} developed by Haehn \etal~\cite{haehn_dojo_2014,Neuroblocks}. Mojo provides a simple scribble interface for error correction, and Dojo extends this for distributed proofreading via a minimalistic web-based user interface. The authors define requirements for general proofreading tools, and then evaluate the accuracy and speed of Raveler, Mojo, and Dojo through a quantitative user study (Sec. 3 and 4)~\cite{haehn_dojo_2014}. In this paper, we use the Dojo system as a baseline for interactive proofreading and extend the experiment reported by Haehn \etal, where Raveler, Mojo, and Dojo are compared in terms of accuracy and speed.
All interactive proofreading solutions require the user to find potential errors manually which takes the majority of time~\cite{proofreading_bottleneck,haehn_dojo_2014}. Recent works propose computer-aided proofreading systems which help with this visual search task.

%Most relevant methods here use heuristics to analyze the image data to find potential errors. 
\textbf{Computer-aided Proofreading.} To reduce the time spent looking for errors, Plaza proposed \textit{focused proofreading} (FP) ~\cite{focused_proofreading}. His approach finds split errors by analyzing segment size ratios across slices and then offers yes/no questions to correct these errors. Plaza reports that additional processing beyond FP is required to find merge errors. His method is freely available as open source software and is integrated into Raveler. This makes it feasible for us to use FP as a baseline for evaluating guided proofreading as described in section \ref{sec:evaluation}. A similar approach was published by Karimov \etal as guided volume editing~\cite{karimov_guided_volume_editing}. Measuring differences in histogram distributions in image data enables to find potential split and merge errors in the corresponding segmentation. For merge errors, the authors generate possible boundaries using watershed which inspired our approach as described in section \ref{sec:methods}. Guided volume editing was designed to let expert users correct labeled computer-tomography datasets by performing several interactions per correction. While focused proofreading and guided volume editing both use a heuristical approach to analyze the image data, Uzunbas \etal showed that potential labeling errors can be found by considering the merge tree of an automatic segmentation method~\cite{uzunbas}. The authors track uncertainty throughout the automatic labeling by training a conditional random field. This method is really a segmentation technique but it is possible to use the uncertainty information to present potential regions for proofreading. Uzunbas' paper describes a great idea but requires further work to overcome the requirement of isotropic volumes, a property not given for most connectomics datasets. Our approach, guided proofreading, works on isotropic as well as anisotropic data, and finds merge and split errors.

>>>>

@@

NeuroProof \cite{neuroproof2013} tries to decrease error rates with interactive learning of agglomeration of over-segmentations of images, based on a random forest classifier. Vazquez-Reina \etal~\cite{amelio_segmentation} propose automatic 3D segmentation by taking whole EM volumes into account rather than a per section approach, then solving a fusion problem with a global context. Kaynig \etal~\cite{kaynig10} propose a random forest classifier coupled with an anisotropic smoothing prior in a conditional random field framework with 3D segment fusion. It is also possible to learn segmentation classification features directly from images with CNNs. Ronneberger \etal~\cite{RonnebergerFB15} use a contracting/expanding CNN path architecture to enable precise boundary localization with small amounts of training data. Lee \etal~\cite{lee2015recursive} recursively train very deep networks with 2D and 3D filters to detect boundaries.

Bogovic \etal~\cite{BogovicHJ13} learn 3D features, and show even that unsupervised learning can produce better features than hand-designs. Our work was inspired by this paper and we extend the features reported by Bogovic \etal for our guided proofreading classifiers as described in section \ref{sec:methods}.
These approaches make good progress; however, in general, proofreading is required to correct errors.

\textbf{Interactive Proofreading.} While proofreading is very time consuming, it is fairly easy for humans to perform corrections through splitting and merging segments. One way to perform such corrections is by using expert tools such as Raveler introduced by Chklovskii~\etal~\cite{chklovskii2010, raveler}. This software offers many parameters for tweaking the proofreading process. Created in 2010, Raveler is still used today by professional full-time proofreaders. Many similar systems exist as stand-alone products or plugins to existing visualization system,~\eg V3D~\cite{proofreading_bottleneck} or AVIZO~\cite{markus_proofreading}.

In contrast to these expert tools, recent works attack the problem of proofreading massive datasets by novices through crowd-sourcing~\cite{saalfeld09,anderson2011,Giuly2013DP2}. A very popular platform is EyeWire presented by Kim \etal~\cite{eyewire_nature}. EyeWire is set up as an online game and participants earn virtual rewards for merging oversegmented labeling to reconstruct the retina cells.
A range of proofreading tools exist in-between expert systems and online games such as Mojo and \textit{Dojo} developed by Haehn \etal~\cite{haehn_dojo_2014,Neuroblocks}. Mojo provides a simple scribble interface for error correction, and Dojo extends this for distributed proofreading via a minimalistic web-based user interface. The authors define requirements for general proofreading tools, and then evaluate the accuracy and speed of Raveler, Mojo, and Dojo through a quantitative user study (Sec. 3 and 4)~\cite{haehn_dojo_2014}. In this paper, we use the Dojo system as a baseline for interactive proofreading and extend the experiment reported by Haehn \etal, where Raveler, Mojo, and Dojo are compared in terms of accuracy and speed.
All interactive proofreading solutions require the user to find potential errors manually which takes the majority of time~\cite{proofreading_bottleneck,haehn_dojo_2014}. Recent works propose computer-aided proofreading systems which help with this visual search task.

%Most relevant methods here use heuristics to analyze the image data to find potential errors. 
\textbf{Computer-aided Proofreading.} To reduce the time spent looking for errors, Plaza proposed \textit{focused proofreading} (FP) ~\cite{focused_proofreading}. His approach finds split errors by analyzing segment size ratios across slices and then offers yes/no questions to correct these errors. Plaza reports that additional processing beyond FP is required to find merge errors. His method is freely available as open source software and is integrated into Raveler. This makes it feasible for us to use FP as a baseline for evaluating guided proofreading as described in section \ref{sec:evaluation}.

A similar approach was published by Karimov \etal as guided volume editing~\cite{karimov_guided_volume_editing}. Measuring differences in histogram distributions in image data enables to find potential split and merge errors in the corresponding segmentation. For merge errors, the authors generate possible boundaries using watershed which inspired our approach as described in section \ref{sec:methods}. Guided volume editing was designed to let expert users correct labeled computer-tomography datasets by performing several interactions per correction.

While focused proofreading and guided volume editing both use a heuristical approach to analyze the image data, Uzunbas \etal showed that potential labeling errors can be found by considering the merge tree of an automatic segmentation method~\cite{uzunbas}. The authors track uncertainty throughout the automatic labeling by training a conditional random field. This method is really a segmentation technique but it is possible to use the uncertainty information to present potential regions for proofreading. Their method requires further work to overcome the requirement of isotropic volumes, a property not given for most connectomics datasets. Our approach, guided proofreading, works on isotropic as well as anisotropic data, and finds merge and split errors.
